Investigating soundscapes perception through acoustic scenes simulation.

### 2017-12-21 lafaySimulatedScenePerception.tex
more

\item \emph{analysis of the semantic features}: the goal is to evaluate if specific classes are more frequently used to simulate a given type of environment (i or ni). To do so, a V-test is considered \citep{lebart2000statistique} \ml{ref en Anglais ?}. With $c$ being the total number of classes used for both types of environment, $c_k$ the number of classes used for a given type of scene $k$ ($k=\lbrace i,ni\rbrace$), $c_j$ the number of time a class $j$ has been used, and $c_{jk}$ the number of time a class $j$ has been used for a given type of scene $k$, the V-test evaluates the null hypothesis that the ratio $\frac{c_{jk}}{c}$ is not significantly different from the ratio $\frac{c_{jk}}{c_k}$. For each class $j$, and each environment type $k$, an approximation of the statistic value $V_{jk}$ is computed as follows:

@@

Concerning $ET$, the $p@5$ is $76\%$ for the abstraction level $0$, and remains above $86\%$ for higher abstraction levels. Considering only the presence / absence of sound classes thus allows us to nicely discriminate between the two types of environments. We also notice that the higher the abstraction level is, thus the more precise the description, the more efective the description is.

@@

\item \emph{differentiating the i- and ni-scenes}: the semantic features, and the global sound levels ($L$, $L(E)$ et $L(T)$), allow us to differentiate between i- and ni-scenes. The semantic description seems to be more powerful;
\item \emph{events or textures}: whatever the feature type, be it semantic or related to the sound pressure level, events are mainly useful to differentiate the two types of scenes; textures bring a limited amount of information;
\item \emph{pleasantness prediction}: considering the correlation between sound levels and pleasantness, it seems that the way subjects perceived the quality of a given environment depends on its nature (i or ni). From the data gathered in those experiments, it is not conceivable to consider the same set of features to predict the pleasantness of i- and ni-scenes in the same time:

@@

\item \emph{for ni-scenes}, global levels ($L$ et $L(E)$), or level of sound markers ($L_m$ et $L(E)_{m}$), have a negative influence on pleasantness. We can notice here that taking into account the contribution of each of the different sources of the scene does not improve the prediction performance compared to a global analysis of the environment.

@@

\item \emph{for i-scenes}, on contrary, the pleasantness prediction needs to consider separately the sound markers characteristics and those of the other sounds. Thus, the markers level, relative to the noise ($L(E)_m-L(E)_b$ and $L_m-L_b$) is positively correlated to the pleasantness, whereas the noise level ($L_b$ and $L(E)_b$) is negatively correlated.

@@

The fact that the pleasantness of the i-scenes is not correlated to global physical features, contrary to pleasantness of ni-scenes, has been studied recently \cite{gozalo2015relationship}.

@@

We observe two perceptual modes of operation that involve different types of features, and that rely on the hedonic nature  of the stimuli. A similar phenomenon is observed for the perception of textures, see Section~\ref{sec:simscene_sampleDataSet}. It seems that the brain adapts its way to encode the information (statistic summary for textures, fine description for events) following a previous decision making with regards to the nature of the stimuli, \ie~is it an event or a texture ? Similarly, in the light of this study, it appears that the features considered for the pleasantness judgment also depends on a preliminary judgment of the global hedonic nature of the environment (ideal or non ideal).

@@

In order to verify that the pleasantness of the ideal and non ideal scenes is affected by the sound markers, the audio waveform of the simulated scenes are regenerated with or without the classes identified as markers. For the i-scenes, some of the i-markers are removed. For the ni-scenes, some of the ni-markers are removed. To evaluate the impact on the perception of pleasantness caused by those modifications, a perceptual test is then conducted with a protocol close to the one considered in experiment 1.b.

@@

\begin{itemize}
\item \emph{for the ni-scenes} we hypothesize that the absence of ni-markers will \textbf{increase} the pleasantness score;
\item \emph{for the i-scenes} we hypothesize that the absence of the i-markers will \textbf{decrease} the pleasantness score.
\end{itemize}

>>>>

\item \emph{analysis of the semantic features}: the goal is to evaluate if specific classes are more frequently used to simulate a given type of environment (i or ni). To do so, a V-test is considered \citep{rakotomalala2008tvpercent}. With $c$ being the total number of classes used for both types of environment, $c_k$ the number of classes used for a given type of scene $k$ ($k=\lbrace i,ni\rbrace$), $c_j$ the number of time a class $j$ has been used, and $c_{jk}$ the number of time a class $j$ has been used for a given type of scene $k$, the V-test evaluates the null hypothesis that the ratio $\frac{c_{jk}}{c}$ is not significantly different from the ratio $\frac{c_{jk}}{c_k}$. For each class $j$, and each environment type $k$, an approximation of the statistic value $V_{jk}$ is computed as follows:

@@

Concerning $ET$, the $p@5$ is $76\%$ for the abstraction level $0$, and remains above $86\%$ for higher abstraction levels. Considering only the presence / absence of sound classes thus allows us to nicely discriminate between the two types of environments. We also notice that the higher the abstraction level is, thus the more precise the description, the more effective the description is.

@@

\item \emph{differentiating the i- and ni-scenes}: the semantic features, and the global sound levels ($L$, $L(E)$ et $L(T)$), allow us to differentiate reliably between i- and ni-scenes. The semantic description seems to be more powerful;
\item \emph{events or textures}: whatever the feature type, be it semantic or related to the sound pressure level, events are the most useful components of the scene to differentiate the two types of scenes; textures bring a limited amount of information;
\item \emph{pleasantness prediction}: considering the correlation between sound levels and pleasantness, it seems that the way subjects perceived the quality of a given environment depends on its very nature (i or ni). From the data gathered in those experiments, the same set of features cannot be considered to predict the pleasantness of i- and ni-scenes:

@@

\item \emph{for ni-scenes}, the global levels ($L$ et $L(E)$), or the level of sound markers ($L_m$ et $L(E)_{m}$), have a negative influence on pleasantness. Taking into account the contribution of each of the different sources of the scene does not improve the prediction performance compared to a global analysis of the environment.

@@

\item \emph{for i-scenes}, on contrary, the sound markers characteristics and those of the other sounds have to be considered separately to predict  the pleasantness. The markers level, relative to the background noise ($L(E)_m-L(E)_b$ and $L_m-L_b$) is positively correlated to the pleasantness, whereas the noise level ($L_b$ and $L(E)_b$) is negatively correlated.

@@

The fact that the pleasantness of the i-scenes is not correlated to global physical features, contrary to pleasantness of ni-scenes, has also been studied recently \cite{gozalo2015relationship}.

@@

We can assume two perceptual modes of operation that involve different types of features and rely on the hedonic nature  of the stimuli. It thus appears that the features considered for the pleasantness judgment also depends on a preliminary judgment of the global hedonic nature of the environment (ideal or non ideal).

A similar phenomenon is observed for the perception of textures, see Section~\ref{sec:simscene_sampleDataSet}. It seems that the brain selectively adapts the way it encodes the information (statistic summary for textures, finer description for events) following a previous decision making process based on the nature of the stimuli, \ie~is it an event or a texture ?

@@

In order to investigat deeper into the relation between the pleasantness of the ideal and non ideal scenes and the sound markers, the audio waveform of the simulated scenes are regenerated with or without the classes identified as markers. To do so, some of the i-markers are removed for the i-scenes and some of the ni-markers are removed for the ni-scenes. To evaluate the impact on the perception of pleasantness caused by those modifications, a perceptual test is conducted with a protocol close to the one considered in experiment 1.b.

@@

\begin{itemize}
\item \emph{for the ni-scenes}, we hypothesize that the absence of ni-markers will \textbf{increase} the pleasantness score;
\item \emph{for the i-scenes}, we hypothesize that the absence of the i-markers will \textbf{decrease} the pleasantness score.
\end{itemize}
